---
title: "Proteomics analysis VS multi-omics CUD"
subtitle: "Version 1"
author: "Frank Stein (PCF) and Eric Zillich"
date: "`r Sys.Date()`"
output: pdf_document
---
 
```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, cache = TRUE)
```
 
```{r echo = FALSE}
setwd("/path/to/protein/")
```
 
This analysis pipeline is partly based on an analysis workflow developed by Bernd Klaus.
 
# R setup

## Defining a working directory

We defined the working directory for the analysis. This directory should contain the metadata file and the tab delimited text files containing the data.

## Loading packages

```{r LoadingPackages}
library(vsn)
library(limma)
library(MSnbase)
library(gplots)
library(fdrtool)
library(biobroom)
library(tidyverse)
```

## Defining some functions and parameters

```{r DefiningFunctions}
customPlot <- list(
  theme_bw(base_size = 12), 
  scale_shape_manual(values = 1:40), 
  scale_fill_brewer(palette = "Set1"), 
  scale_colour_brewer(palette = "Set1")
)
script.version <- "V1"
dir_save <- paste0("data_analysis_results_", script.version)
if (!dir.exists(dir_save))
{
  dir.create(dir_save)
}
```

# Loading data and annotating experimental conditions

```{r loadRawFiles}
conditions <- read.delim("metadata_detailed.csv", sep = ",")
files <- file.path(unique(conditions$file))
files <- files[file.exists(files)]
data <- NULL
for (i in seq_along(files))
{
  print(files[i])
  x <- read_delim(file.path(files[i]), delim = "	")
  names(x) <- make.names(names(x))
  x$Gene[is.na(x$Gene)] <- x$Protein.ID[is.na(x$Gene)]
  #keep only proteins with at least two unique peptide matches
  x <- x %>%
    dplyr::filter(Razor.Peptides >= 2)
  #remove reverse hits
  x <- x %>%
    dplyr::filter(!grepl("rev_", Protein))
  #remove known contaminants
  x <- x %>%
    dplyr::filter(!grepl("contam_", Protein))
  x$file <- files[i]
  keepnames <- as.character(subset(conditions, file == files[i])$col.name)
  x <- x[, c("Gene", "Protein.ID", "Protein.Description", "Organism", "Unique.Peptides", "Razor.Peptides", "Total.Intensity", "file", keepnames)]
  x <- x %>%
    group_by(Gene, Protein.ID, Protein.Description, Organism, Unique.Peptides, Razor.Peptides, Total.Intensity, file) %>%
    gather(key = "col.name", value = "value", keepnames)
  x$Total.Intensity <- as.numeric(as.character(x$Total.Intensity))
  data <- bind_rows(data, x)
  rm(x, keepnames)
}
rm(i, files)
data <- left_join(data, conditions)
data <- subset(data, value > 0)
```

# Protein identification overview

```{r proteinIdentificationOverview}
sub <- unique(data[, c("Gene", "file", "rep")])
sub$found <- 1
sub$file_old <- sub$file
sub$file <- paste("file", as.numeric(factor(sub$file)))
print(unique(with(sub, paste(file, "-", file_old))))
sub$file_old <- NULL
sub <- unique(sub)
ggplot(data = sub, aes(file, fill = rep)) + 
  geom_bar(position = position_dodge()) + 
  theme_bw(base_size = 12) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
with(sub, table(file, rep))
sub_i <- data %>%
  group_by(Gene) %>%
  dplyr::select(Gene, rep, file) %>%
  unique() %>%
  summarise(found.in.files = n())
ggplot(data = sub_i, aes(found.in.files)) + 
  geom_bar(position = position_dodge()) + 
  customPlot + 
  xlab("no of identifications per file")
table(sub_i$found.in.files)
rm(sub_i)
rm(sub)
```

# Identification and removal of duplicated gene names

```{r renameDuplicates}
dup.data <- data %>%
  ungroup() %>%
  dplyr::select(Gene, Protein.ID, file, Razor.Peptides) %>%
  unique() %>%
  group_by(file, Gene) %>%
  dplyr::filter(n() > 1)
if (nrow(dup.data) > 0)
{
  write.csv(dup.data, file.path(dir_save, paste0("Duplicates_", script.version, ".csv")), row.names = FALSE)
}
dups <- unique(dup.data$Gene)
if (length(dups) > 0)
{
  print("Duplicates were identified:")
  print(dups)
  print("Removing genes with smaller numbers of quantified peptides.")
  for (dup in dups)
  {
    for (filen in unique(subset(dup.data, Gene == dup)$file))
    {
      sub <- subset(data, Gene == dup & file == filen)
      data <- subset(data, paste(Gene, file) != paste(dup, filen))
      sub <- subset(sub, Razor.Peptides == max(sub$Razor.Peptides, na.rm = TRUE))
      data <- bind_rows(data, sub)
      rm(sub)
    }
    rm(filen)
  }
  rm(dup)
}
dup.data <- data %>%
  ungroup() %>%
  dplyr::select(Gene, Protein.ID, file, Razor.Peptides) %>%
  unique() %>%
  group_by(file, Gene) %>%
  dplyr::filter(n() > 1)
dups <- unique(dup.data$Gene)
if (length(dups) > 0)
{
  print("Duplicates:")
  print(dups)
  print("Removing genes with the least overlap of protein ids in other ms experiments")
  for (dup in dups)
  {
    sub <- subset(data, Gene %in% dup)
    IDs <- unlist(strsplit(sub$Protein.ID, split = "[|]"))
    rm(sub)
    for (filen in unique(subset(dup.data, Gene == dup)$file))
    {
      sub <- subset(data, Gene == dup & file == filen)
      data <- subset(data, paste(Gene, file) != paste(dup, filen))
      sub$ID.count <- 0
      for (ID in unique(sub$Protein.ID))
      {
        sub$ID.count[sub$Protein.ID == ID] <- length(which(IDs %in% unlist(strsplit(ID, split = "[|]"))))
      }
      sub <- subset(sub, ID.count == max(sub$ID.count, na.rm = TRUE))
      sub$ID.count <- NULL
      data <- bind_rows(data, sub)
      rm(sub)
    }
    rm(filen, IDs)
  }
  rm(dup)
}
rm(dups, dup.data)
```

# Transforming long data to wide data

```{r transformLongData2WideData, fig.height = 5}
ggplot(data = data, aes(condition, log2(value), fill = batch)) + 
  geom_boxplot(aes(group = paste(batch, condition, rep))) + 
  ylab("log2(reporter_intensity)") + 
  customPlot + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
cdata <- data %>%
  group_by(Gene) %>%
  mutate(key = paste("reporter_intensity", condition, rep, sep = "_")) %>%
  dplyr::select(Gene, key, value) %>%
  group_by(Gene, key) %>%
  summarise(value = sum(value, na.rm = TRUE)) %>%
  spread(key = key, value = value)
#found.in.files
fdata <- data %>%
  group_by(Gene) %>%
  dplyr::select(Gene, file) %>%
  unique() %>%
  summarise(found.in.files = n())
fdata_i <- data %>%
  mutate(file2 = make.names(file)) %>%
  group_by(Gene, file2) %>%
  dplyr::select(Gene, file, file2) %>%
  unique() %>%
  summarise(found.in.file = n()) %>%
  mutate(file2 = paste0("found.in.file_", file2)) %>%
  spread(key = file2, value = found.in.file)
fdata <- left_join(fdata, fdata_i)
#found.in.conditions
fdata_i <- data %>%
  group_by(Gene) %>%
  dplyr::select(Gene, condition) %>%
  unique() %>%
  summarise(found.in.conditions = n())
fdata <- left_join(fdata, fdata_i)
#found.in.reps
fdata_i <- data %>%
  group_by(Gene) %>%
  dplyr::select(Gene, rep) %>%
  unique() %>%
  summarise(found.in.reps = n())
fdata <- left_join(fdata, fdata_i)
#Unique.Peptides overview
fdata_i <- data %>%
  group_by(Gene) %>%
  dplyr::select(Gene, Unique.Peptides) %>%
  unique() %>%
  summarise(max.Unique.Peptides = max(Unique.Peptides, na.rm = TRUE))
fdata <- left_join(fdata, fdata_i)
#Razor.Peptides overview
fdata_i <- data %>%
  group_by(Gene) %>%
  dplyr::select(Gene, Razor.Peptides) %>%
  unique() %>%
  summarise(max.Razor.Peptides = max(Razor.Peptides, na.rm = TRUE))
fdata <- left_join(fdata, fdata_i)
#abundance overview
fdata_i <- data %>%
  group_by(Gene) %>%
  dplyr::select(Gene, Total.Intensity) %>%
  unique() %>%
  summarise(average.Total.Intensity = mean(Total.Intensity, na.rm = TRUE))
fdata <- left_join(fdata, fdata_i)
id_data <- data %>%
  ungroup() %>%
  dplyr::select(Gene, Protein.ID, Protein.Description, Organism) %>%
  unique()
dups <- id_data %>%
  group_by(Gene) %>%
  summarize(n = n()) %>%
  dplyr::filter(n > 1) %>%
  dplyr::select(Gene)
if (nrow(dups) > 0)
{
  dups <- dups$Gene
  dup.data <- subset(id_data, Gene %in% dups)
  id_data <- subset(id_data, !Gene %in% dups)
  combine_ids <- function(ids)
  {
    ids <- as.character(ids)
    return(paste(
      sort(unique(unlist(strsplit(ids, split = "[;]")))), collapse = ";"))
  }
  dup.data <- dup.data %>%
    group_by(Gene) %>%
    summarise(Protein.ID = combine_ids(Protein.ID), 
              Protein.Description = combine_ids(Protein.Description), 
              Organism = combine_ids(Organism))
  id_data <- bind_rows(id_data, dup.data)
  rm(dup.data, combine_ids)
}
rm(dups)
fdata <- left_join(id_data, fdata)
rm(id_data)
cdata <- left_join(fdata, cdata)
rm(fdata_i, fdata)
```

# Filter data

Only proteins that were quantified with two unique peptide matches are kept for the analysis. Proteins were filtered according to these condition already when loading the data. Moreover, only proteins were kept if they were quantified in at least 2/3 of the replicates.

```{r filterData}
dim(cdata)
min.found.in.files <- 3
min.Razor.Peptides <- 2
cdata <- cdata %>%
  dplyr::filter(found.in.files >= min.found.in.files, max.Razor.Peptides >= min.Razor.Peptides)
dim(cdata)
rm(min.found.in.files, min.Razor.Peptides)
```

# Building an expression set object

```{r buildingEset}
# Constructing assay data
raw_data <- cdata %>%
  dplyr::select(starts_with("reporter_intensity")) %>%
  as.data.frame()
rownames(raw_data) <- cdata$Gene
names(raw_data) <- gsub("reporter_intensity_", "", names(raw_data))

# Constructing metadata
conditions_i <- data.frame(ID = names(raw_data))
conditions_i <- left_join(conditions_i, conditions)

# Constructing fdata
fdata <- cdata %>%
  dplyr::select(-starts_with("reporter_intensity")) %>%
  as.data.frame()

# Defining ID columns
rownames(fdata) <- fdata$Gene
rownames(conditions_i) <- conditions_i$ID
colnames(raw_data) <- conditions_i$ID

# Log2-transformation of raw_ reporter_intensitys
raw_data_m <- log2(as.matrix(raw_data))
raw_data_m[is.infinite((raw_data_m))] <- NA
raw_data_m[is.na((raw_data_m))] <- NA

# Creating an expression set
raw_dataE <- ExpressionSet(assayData = raw_data_m, 
                           phenoData = AnnotatedDataFrame(conditions_i), 
                           featureData = AnnotatedDataFrame(fdata))
validObject(raw_dataE)
rm(raw_data, raw_data_m, conditions_i, fdata)
```

# Batch effect removal

```{r removingBatchEffects}
batchcl_dataE <- raw_dataE
exprs(batchcl_dataE) <- removeBatchEffect(exprs(batchcl_dataE), 
      batch = as.character(pData(batchcl_dataE)$batch), 
      design = model.matrix(~0 + as.character(pData(batchcl_dataE)$condition)))
```

# Data normalization

The vsn package from Wolfgang Huber is used to apply a variance stabilization normalization method on the log2 raw data.

```{r DataNormalization}
norm_dataE <- batchcl_dataE
vsn.fit <- vsn2(2^exprs(batchcl_dataE))
meanSdPlot(vsn.fit)
norm_dataE <- batchcl_dataE
exprs(norm_dataE) <- predict(vsn.fit, 2^exprs(norm_dataE))

rm(sdplot.object, vsn.fit)
```

# Calculation of ctrl-fold changes

```{r CalculateCtrlFoldChanges}
fc.data <- as.data.frame(2^exprs(norm_dataE))
names.orig <- names(fc.data)
for (i in seq_along(conditions$ID))
{
  fc.data[, paste0(conditions$ID[i], ".ctrl.ratio")] <-
    fc.data[, as.character(conditions$ID[i])] /
    apply(fc.data[, subset(conditions, condition == "CTRL" & batch == conditions$batch[i])$ID], 1, median, na.rm = TRUE)
    # fc.data[, as.character(conditions$ctrl.ID[i])]
}
rm(i)
fc.data <- fc.data %>%
  dplyr::select(ends_with(".ctrl.ratio"))
fc.data$Gene <- rownames(fc.data)
fc.data <- fc.data %>%
  dplyr::select(ends_with(".ctrl.ratio"))
ctrl.ratio_dataE <- norm_dataE
names(fc.data) <- gsub(".ctrl.ratio", "", names(fc.data))
fc.data <- fc.data[ , colnames(ctrl.ratio_dataE)]
exprs(ctrl.ratio_dataE) <- fc.data %>%
  as.matrix() %>%
  log2()
rm(fc.data)
```

# PCA - principal component analysis 

```{r CreatePCAplots, fig.height = 5}
sets <- list("raw data" = raw_dataE, "batchcl data" = batchcl_dataE, "norm data" = norm_dataE, "ctrl.ratio data" = ctrl.ratio_dataE)
PCA_data <- NULL
for (i in seq_along(sets)) {
  set <- sets[[i]]
  set.name <- names(sets)[i]
  PCA_m <- t(na.omit(exprs(set)))
  if (length(PCA_m) > 50) {
    PCA <- prcomp(PCA_m, scale = FALSE)
    perc_var <-
      round(100 * PCA$sdev ^ 2 /
              sum(PCA$sdev ^ 2), 1)
    PCA_data_i <-
      data.frame(PC1 = PCA$x[, 1],
                 PC2 = PCA$x[, 2],
                 PC3 = PCA$x[, 3],
                 PC1.var = perc_var[1],
                 PC2.var = perc_var[2],
                 condition = pData(set)$condition,
                 rep = pData(set)$batch,
                 measurement = set.name)
    PCA_data <- bind_rows(PCA_data, PCA_data_i)
    rm(PCA_data_i)
  }
}
rm(i)
PCA_data$measurement <- factor(PCA_data$measurement, ordered = TRUE, levels = c("raw data", "batchcl data", "norm data", "ctrl.ratio data"))

ggplot(data = PCA_data, aes(PC1, PC2, colour = PC3)) +
  geom_point(aes(colour = condition, shape = rep), size = 4) +
  guides(size = "none") +
  customPlot +
  ggtitle("PCA analysis") +
  xlab("PC1") +
  ylab("PC2") +
  facet_wrap(. ~ measurement + paste("PC1:", PC1.var, "% var - PC2:", PC2.var, "% var"), scale = "free", ncol = 5)
ggsave(file.path(dir_save, paste0("PCA_analysis_", script.version, ".pdf")),
     width = 17, height = 5)
write.csv(PCA_data, file = file.path(dir_save, paste0("PCA_analysis_data_", script.version, ".csv")), row.names = FALSE)

# Additional PCA plot for the supplement 

PCA_data2 <- PCA_data[PCA_data$measurement =="norm data",]
rnames <- substr(rownames(PCA_data2),10,12)
rownames(PCA_data2) <- rnames
# Import additional pheno information
pheno_with_CUD_Ctrl <- read.csv("/path/to/RNA/DE_results/pheno_with_CUD_Ctrl.txt", sep=";")
rownames(pheno_with_CUD_Ctrl) <- pheno_with_CUD_Ctrl$Brain_ID

PCA_data2 <- merge(pheno_with_CUD_Ctrl,PCA_data2,by=0)

# Plot PCA
library(ggrepel)
pdf("/path/to/protein/data_analysis_results_V1/PCA_Proteomics_Supp.pdf", width = 5, height = 5 )
ggplot(data=PCA_data2, aes_string(x="PC1", y="PC2", color="condition")) + geom_point(size=1) + 
  xlab(paste0("PC1: ",round(PCA_data2$PC1.var[1]),"% variance")) +ylim(-60,60)+xlim(-60,60)+
  ylab(paste0("PC2: ",round(PCA_data2$PC2.var[1]),"% variance")) +
  coord_fixed(ratio=1)+ geom_text_repel(size = 3, aes(label = CUD), vjust = 2.2,max.overlaps = 3) +theme_minimal()+scale_color_manual(values=c("black","black"))
dev.off()

```

# Merge modified data back into 'cdata'

```{r CreateCastedData}
cdata_i <- as.data.frame(2^exprs(batchcl_dataE))
names(cdata_i) <- paste0("batchcl_reporter_intensity_", names(cdata_i))
cdata_i$Gene <- rownames(cdata_i)
cdata <- left_join(cdata, cdata_i)
rm(cdata_i)

cdata_i <- as.data.frame(2^exprs(norm_dataE))
names(cdata_i) <- paste0("norm_reporter_intensity_", names(cdata_i))
cdata_i$Gene <- rownames(cdata_i)
cdata <- left_join(cdata, cdata_i)
rm(cdata_i)

cdata_i <- as.data.frame(2^exprs(ctrl.ratio_dataE))
names(cdata_i) <- paste0("ctrl.ratio_", names(cdata_i))
cdata_i$Gene <- rownames(cdata_i)
cdata <- left_join(cdata, cdata_i)
rm(cdata_i)

write.csv(cdata, file = file.path(dir_save, paste0("Full_dataset_", script.version, ".csv")), row.names = FALSE)
```

# Create tidy data

```{r CreateTidyData}
mdata <- NULL

mdata_i <- tidy(raw_dataE)
mdata_i <- mdata_i %>%
  mutate(value = 2 ^ value)
names(mdata_i)[1] <- "Gene"
names(mdata_i)[2] <- "ID"
mdata_i <- left_join(mdata_i, conditions)
mdata_i$measurement <- "raw_reporter_intensity"
mdata <- bind_rows(mdata, mdata_i)
rm(mdata_i)

mdata_i <- tidy(batchcl_dataE)
mdata_i <- mdata_i %>%
  mutate(value = 2 ^ value)
names(mdata_i)[1] <- "Gene"
names(mdata_i)[2] <- "ID"
mdata_i <- left_join(mdata_i, conditions)
mdata_i$measurement <- "batchcl_reporter_intensity"
mdata <- bind_rows(mdata, mdata_i)
rm(mdata_i)

mdata_i <- tidy(norm_dataE)
mdata_i <- mdata_i %>%
  mutate(value = 2 ^ value)
names(mdata_i)[1] <- "Gene"
names(mdata_i)[2] <- "ID"
mdata_i <- left_join(mdata_i, conditions)
mdata_i$measurement <- "norm_reporter_intensity"
mdata <- bind_rows(mdata, mdata_i)
rm(mdata_i)

mdata_i <- tidy(ctrl.ratio_dataE)
mdata_i <- mdata_i %>%
  mutate(value = 2 ^ value)
names(mdata_i)[1] <- "Gene"
names(mdata_i)[2] <- "ID"
mdata_i <- left_join(mdata_i, conditions)
mdata_i$measurement <- "ctrl.ratio"
mdata <- bind_rows(mdata, mdata_i)
rm(mdata_i)

mdata$condition <- factor(mdata$condition, ordered = TRUE, levels = c("CASE", "CTRL"))
mdata$measurement <- factor(mdata$measurement, ordered = TRUE, levels = c("raw_reporter_intensity", "batchcl_reporter_intensity", "norm_reporter_intensity", "ctrl.ratio"))
```

# Data transformation overview

```{r DataTransformationOverview, fig.height = 6}
customPlot <- list(
  theme_bw(base_size = 12), 
  scale_shape_manual(values = 1:40)
)
library(RColorBrewer)
mycolors <- colorRampPalette(brewer.pal(8, "Set2"))(40)

ggplot(data = subset(mdata, !grepl("ratio", measurement)), aes(condition, log2(value))) +
  geom_boxplot(aes(fill = rep)) +scale_fill_manual(values=mycolors)+
  customPlot +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  facet_grid(. ~ measurement)
ggsave(file.path(dir_save, paste0("Normalization_overview_", script.version, ".pdf")), width = 12, height = 5)

ggplot(data = subset(mdata, grepl("ratio", measurement)), aes(condition, log2(value))) +
  geom_boxplot(aes(fill = rep)) +scale_fill_manual(values=mycolors)+
  customPlot +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  facet_grid(. ~ measurement, scale = "free")
ggsave(file.path(dir_save, paste0("Normalization_overview_ratios_", script.version, ".pdf")), width = 6, height = 5)
```

# LIMMA analysis
## Initialize limma results data

```{r LimmaInitialization}
limma_results <- NULL
```

## Run limma analysis

```{r LimmaRun}
# Export the different datasets
write.table(exprs(raw_dataE),"/path/to/protein/data_analysis_results_V1/raw_data.txt",sep=";",quote=F,row.names = T,col.names = T)
write.table(exprs(norm_dataE),"/path/to/protein/data_analysis_results_V1/normalized_data.txt",sep=";",quote=F,row.names = T,col.names = T)
write.table(exprs(batchcl_dataE),"/path/to/protein/data_analysis_results_V1/batch_corrected_data.txt",sep=";",quote=F,row.names = T,col.names = T)

# remove PCA outliers
limma_data <- norm_dataE[, !(sampleNames(norm_dataE) %in% c("CASE_rep_ID1_1","CASE_rep_ID2"))]
limma_weights <- exprs(raw_dataE[, !(sampleNames(raw_dataE) %in% c("CASE_rep_ID1_1","CASE_rep_ID2"))])
comparison.list <- c("CASE - CTRL")
limma_weights <- ifelse(is.na(limma_weights), 0.01, 1)

# Add further covariate data to the limma analysis
pheno <- pData(limma_data)
pheno$SampleID <- gsub("_","",pheno$SampleID)

# Import CUD phenotype data and add to pData
pheno_CUD_all <- read.csv2("/path/to/pheno/CUD_randomised_20220707_layout.csv")
colnames(pheno_CUD_all)[1]<-"SampleID"
pheno_merged <- merge(pheno, pheno_CUD_all,by="SampleID")
pheno_merged$pH[is.na(pheno_merged$pH)] <- 6.335
rownames(pheno_merged) <- pheno_merged$ID
pheno_merged <- pheno_merged[order(as.numeric(pheno_merged$SampleID),decreasing=F),]
pheno_merged <- pheno_merged[match(rownames(pheno),rownames(pheno_merged)),]
write.table(pheno_merged,"/path/to/protein/data_analysis_results_V1/pheno_merged_with_proteomics.txt",sep=";",quote=F,row.names = T,col.names = T)

limma.PMI <- as.numeric(pheno_merged$PMI)
limma.pH <- as.numeric(pheno_merged$pH)
limma.Age <- as.numeric(pheno_merged$Age)
limma.batch <- factor(pheno_merged$batch,ordered = FALSE)
limma.cond <- factor(pheno_merged$condition, ordered = FALSE)

contrast.matrix <- model.matrix( ~  0 + limma.cond + limma.batch + limma.pH + limma.PMI + limma.Age)
colnames(contrast.matrix) <- gsub("limma.cond", "", colnames(contrast.matrix))

limma.object <- eBayes(
  contrasts.fit(
    lmFit(limma_data, design = contrast.matrix, weights = limma_weights), 
    makeContrasts(contrasts = comparison.list, levels = contrast.matrix)
  )
)
rm(comparison.list, comparison, contrast.matrix, limma.cond, limma.rep, limma_data, limma_weights, limma.object)
```

## Volcano-plot

```{r LimmaVolcanoPlot, fig.height = 7}
limma_results$DE <- 
  with(limma_results, ifelse(pvalue.limma < 0.05 & logFC > 0.07, "DE up p<0.05",
                                 ifelse(pvalue.limma < 0.05 & logFC < -0.07, "DE down p<0.05", "n.s.")))
limma_results$DE <- 
  factor(limma_results$DE, 
         ordered = TRUE, 
         levels = c("DE up p<0.05","DE down p<0.05","n.s."))

customPlot2 <- list(
  theme_minimal(base_size = 12), 
  scale_fill_manual(values=c("#E43F3F","#268989","gray80")), 
  scale_colour_manual(values=c("#E43F3F","#268989","gray80"))
)

pdf("/path/to/protein/DE_volcano_manual_v1.pdf", width = 10, height = 6)
ggplot(data = limma_results, aes(logFC, -log10(pvalue.limma), colour = DE)) +
  geom_vline(aes(xintercept = 0)) + xlim(-1,1)+
  geom_point() + geom_hline(aes(yintercept = -log10(0.05)),size=0.2) + geom_hline(aes(yintercept = -log10(1.17096e-05)),size=0.2,linetype="dashed")+
  geom_text(aes(label = gsub("[|].+", "", Gene)), 
            data = subset(limma_results, DE != "n.s."), 
            vjust = 0, nudge_y = 0.1, size = 4, check_overlap = TRUE) +
  xlab("log2(fold change)") + ylab("-log10(pvalue)") +
  customPlot2
dev.off()
```

## MA-plot

```{r LimmaMA_plot, fig.height = 7}
ggplot(data = limma_results, aes(AveExpr, logFC, colour = hit_annotation)) +
  geom_hline(aes(yintercept = 0)) + 
  geom_point() +
  geom_text(aes(label = gsub("[|].+", "", Gene)), 
            data = subset(limma_results, hit_annotation != "no hit"), 
            vjust = 0, nudge_y = 0.1, size = 2, check_overlap = TRUE) + 
  facet_wrap( ~ comparison + hit_annotation_method) +
  xlab("average log2(reporter_intensity)") +
  ylab("log2(fold change)") +
  customPlot
ggsave(file.path(dir_save, paste0("MA_plot_", script.version, ".pdf")), width = 7, height = 5)
```

##Total.Intensity-plot

```{r LimmaAbundancePlot, fig.height = 7}
limma_results$Total.Intensity <- limma_results$average.Total.Intensity
ggplot(data = limma_results, aes(log2(Total.Intensity), logFC, colour = hit_annotation)) +
  geom_hline(aes(yintercept = 0)) + 
  geom_point() +
  geom_text(aes(label = gsub("[|].+", "", Gene)), 
            data = subset(limma_results, hit_annotation != "no hit"), 
            vjust = 0, nudge_y = 0.1, size = 2, check_overlap = TRUE) + 
  facet_wrap( ~ comparison + hit_annotation_method) +
  ylab("log2(fold change)") +
  customPlot
ggsave(file.path(dir_save, paste0("Total.Intensity_plot_", script.version, ".pdf")), width = 7, height = 5)
```

## Save limma results

```{r LimmaSave}
write.csv(limma_results, file.path(dir_save, paste0("Limma_results_", script.version, ".csv")), row.names = FALSE)
```

